---
title: "Homework 3"
author: "Shina Min"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output: github_document

---

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)

library(p8105.datasets)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

### Problem 0

This solution focuses on a reproducible report containing code and text necessary for Problems 1-3, and is organized as an R Project. This was not prepared as a GitHub repo; examples for repository structure and git commits should be familiar from other elements of the course.

Throughout, we use appropriate text to describe our code and results, and use clear styling to ensure code is readable. 

### Problem 1

#### Read in the data

```{r}
data("instacart")

instacart = 
  instacart |> 
  as_tibble()
```

#### Answer questions about the data

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns, with each row resprenting a single product from an instacart order. Variables include identifiers for user, order, and product; the order in which each product was added to the cart. There are several order-level variables, describing the day and time of the order, and number of days since prior order. Then there are several item-specific variables, describing the product name (e.g. Yogurt, Avocado), department (e.g. dairy and eggs, produce), and aisle (e.g. yogurt, fresh fruits), and whether the item has been ordered by this user in the past. In total, there are `r instacart |> select(product_id) |> distinct() |> count()` products found in `r instacart |> select(user_id, order_id) |> distinct() |> count()` orders from `r instacart |> select(user_id) |> distinct() |> count()` distinct users.

Below is a table summarizing the number of items ordered from aisle. In total, there are 134 aisles, with fresh vegetables and fresh fruits holding the most items ordered by far.

```{r}
instacart |> 
  count(aisle) |> 
  arrange(desc(n))
```

Next is a plot that shows the number of items ordered in each aisle. Here, aisles are ordered by ascending number of items.

```{r}
instacart |> 
  count(aisle) |> 
  filter(n > 10000) |> 
  mutate(aisle = fct_reorder(aisle, n)) |> 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

```

Our next table shows the three most popular items in aisles `baking ingredients`, `dog food care`, and `packaged vegetables fruits`, and includes the number of times each item is ordered in your table.

```{r}
instacart |> 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |>
  group_by(aisle) |> 
  count(product_name) |> 
  mutate(rank = min_rank(desc(n))) |> 
  filter(rank < 4) |> 
  arrange(desc(n)) |>
  knitr::kable()
```

Finally is a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week. This table has been formatted in an untidy manner for human readers. Pink Lady Apples are generally purchased slightly earlier in the day than Coffee Ice Cream, with the exception of day 5.

```{r}
instacart |>
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |>
  group_by(product_name, order_dow) |>
  summarize(mean_hour = mean(order_hour_of_day)) |>
  pivot_wider(
    names_from = order_dow, 
    values_from = mean_hour) |>
  knitr::kable(digits = 2)
```

### Problem 2


his problem uses the BRFSS data. DO NOT include this dataset in your local data directory; instead, load the data from the p8105.datasets package using


## Loading the package for Problem 2.
```{r}
library(p8105.datasets)
data("brfss_smart2010")
```


## Formatting, organizing, and data cleaning.
```{r}
brffs_smart2010_df=
  brfss_smart2010 %>%
  janitor::clean_names() %>%
  filter(topic == "Overal Health" &
           response %in% c("Excellent", "Very Good", "Good", "Fair", "Poor")) %>%
  mutate(response = factor(response, levels = c("Poor", "Fair", "Good", "Very Good", "Excellent")))
```


## States observed at 7 or more locations in 2002 and 2010.
```{r}
brffs_smart2010_df %>%
  filter(year %in% c(2002, 2010)) %>% 
  group_by(year, locationabbr) %>%
  summarize(n_obs = n()) %>%
  filter(n_obs >= 7) %>%
  arrange(year, desc(n_obs))
```


## Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state.
```{r}
excellent_data =
  brfss_smart2010 %>%
  filter(Response %in% c("Excellent")) %>%
  select(Year, Locationabbr, Data_value) %>%
  group_by(Locationabbr, Year) %>%
  mutate(avg_data_value = mean(Data_value)) %>%
  select(Year, Locationabbr, avg_data_value) %>%
  distinct()
```

* In 2002, PA, MA, NJ, CT, FL, NC (total of 6 states) were observed at 7 or more locations. In 2010, total of 14 states were observed at 7 or more locations, including FL, NJ, TX, CA, MD, NC, NE, WA, MA, NY, OH, CO, PA, SC.
* WA were observed at 7 or more locations in 2010.

## Making a "spaghetti" plot
```{r}
excellent_data %>%
  ggplot(aes(x = Year, y = avg_data_value, color = Locationabbr)) +
  geom_line() +
  labs(title = "Average Value of 'Excellent' Responses Over Time", x = "Year", y = "Average Value") +
  theme_minimal()
```

* Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.

## Making a two-panel plot
```{r}
brfss_smart2010 %>%
  select(Year, Data_value, Locationabbr, Locationdesc, Response) %>%
  filter(
    Year %in% c(2006, 2010),
    Locationabbr == "NY") %>% 
  ggplot(aes(x = Data_value, fill = Response)) +
  geom_boxplot(alpha = 0.5) +
  labs(title = "Distribution of Data_value in NY State (2006 & 2010)",
       x = "Data Value", y = "Response") +
  theme_minimal() +
  facet_grid(. ~ Year)
```



## Problem 3

```{r}
library(dplyr)
```

## Load, tidy, merge and organize the data sets.
Load, tidy, merge, and otherwise organize the data sets. Your final dataset should include all originally observed variables; exclude participants less than 21 years of age, and those with missing demographic data; and encode data with reasonable variable classes (i.e. not numeric, and using factors with the ordering of tables and plots in mind).
```{r}
covar_df= 
  read.csv("~/Desktop/Hw3/nhanes_covar.csv", skip = 4) %>%
  janitor::clean_names() %>%
  drop_na() %>%
  filter(age >= 21) %>% 
  mutate(education = recode(education, "1" = "Less than high school", "2"= "High school equivalent", "3" = "More than high school")) %>%
  mutate(
    sex = recode(sex, "1" = "male", "2" = "female"),
    education = recode(education, "1" = "male")
  )

accel_df =
  read.csv("~/Desktop/Hw3/nhanes_accel.csv") %>%
  janitor::clean_names()

joined_data =
  accel_df %>%
  inner_join(covar_df, by = "seqn") %>%
  relocate(seqn, sex, age, bmi, education) %>%
  filter(age >= 21) %>%
  mutate(
    sex = forcats::fct_relevel(sex, "male", "female"),
    education = forcats::fct_relevel(education, "Less than high school", "High school equivalent", "More than high school")
  )
view(joined_data)
```



## Creating a table for the number of men and women in each category.
ADD COMMENT!
```{r}
education_table =
  joined_data %>%
  group_by(education, sex) %>%
  summarize(count = n())

knitr::kable(education_table, caption = "Number of Men and Women in each Education Category", format = "markdown")
```


## Plotting age_distribution boxplot
Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate across minutes to create a total activity variable for each participant. Plot these total activities (y-axis) against age (x-axis); your plot should compare men to women and have separate panels for each education level. Include a trend line or a smooth to illustrate differences. Comment on your plot.
```{r}
age_distribition_viz =
  ggplot(joined_data, aes(x = education, y = age, fill = sex)) +
  geom_boxplot() +
  labs(title = "Age Distribution by Gender and Education",
       x = "Education Level",
       y = "Age") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


Accelerometer data allows the inspection activity over the course of the day. Make a three-panel plot that shows the 24-hour activity time courses for each education level and use color to indicate sex. Describe in words any patterns or conclusions you can make based on this graph; including smooth trends may help identify differences.









