---
title: "Homework 3"
author: "Shina Min"
date: "2023.10.14"
output: github_document

---

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)

library(p8105.datasets)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

### Problem 0

This solution focuses on a reproducible report containing code and text necessary for Problems 1-3, and is organized as an R Project. This was not prepared as a GitHub repo; examples for repository structure and git commits should be familiar from other elements of the course.

Throughout, we use appropriate text to describe our code and results, and use clear styling to ensure code is readable. 

### Problem 1

#### Read in the data

```{r}
data("instacart")

instacart = 
  instacart |> 
  as_tibble()
```

#### Answer questions about the data

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns, with each row resprenting a single product from an instacart order. Variables include identifiers for user, order, and product; the order in which each product was added to the cart. There are several order-level variables, describing the day and time of the order, and number of days since prior order. Then there are several item-specific variables, describing the product name (e.g. Yogurt, Avocado), department (e.g. dairy and eggs, produce), and aisle (e.g. yogurt, fresh fruits), and whether the item has been ordered by this user in the past. In total, there are `r instacart |> select(product_id) |> distinct() |> count()` products found in `r instacart |> select(user_id, order_id) |> distinct() |> count()` orders from `r instacart |> select(user_id) |> distinct() |> count()` distinct users.

Below is a table summarizing the number of items ordered from aisle. In total, there are 134 aisles, with fresh vegetables and fresh fruits holding the most items ordered by far.

```{r}
instacart |> 
  count(aisle) |> 
  arrange(desc(n))
```

Next is a plot that shows the number of items ordered in each aisle. Here, aisles are ordered by ascending number of items.

```{r}
instacart |> 
  count(aisle) |> 
  filter(n > 10000) |> 
  mutate(aisle = fct_reorder(aisle, n)) |> 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

```

Our next table shows the three most popular items in aisles `baking ingredients`, `dog food care`, and `packaged vegetables fruits`, and includes the number of times each item is ordered in your table.

```{r}
instacart |> 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |>
  group_by(aisle) |> 
  count(product_name) |> 
  mutate(rank = min_rank(desc(n))) |> 
  filter(rank < 4) |> 
  arrange(desc(n)) |>
  knitr::kable()
```

Finally is a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week. This table has been formatted in an untidy manner for human readers. Pink Lady Apples are generally purchased slightly earlier in the day than Coffee Ice Cream, with the exception of day 5.

```{r}
instacart |>
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |>
  group_by(product_name, order_dow) |>
  summarize(mean_hour = mean(order_hour_of_day)) |>
  pivot_wider(
    names_from = order_dow, 
    values_from = mean_hour) |>
  knitr::kable(digits = 2)
```

### Problem 2


## Loading the package for Problem 2.
```{r}
library(p8105.datasets)
data("brfss_smart2010")
```

  * Loaded `library(p8105.datasets)` and `data("brfss_smart2010")` to start the assignment!

## Formatting, organizing, and data cleaning.
```{r}
brffs_smart2010_df=
  brfss_smart2010 %>%
  janitor::clean_names() %>%
  filter(topic == "Overal Health" &
           response %in% c("Excellent", "Very Good", "Good", "Fair", "Poor")) %>%
  mutate(response = factor(response, levels = c("Poor", "Fair", "Good", "Very Good", "Excellent")))
```

  * Defined `brffs_smart2010` by giving it a name `brfss_smart2010_df`.
  * Cleaned the dataset using `janitor::clean_names` function.
  * Then filtered the responses in order from "Poor" to "Excellent" by using `mutate` and `filter` functions.


## States observed at 7 or more locations in 2002 and 2010.
```{r}
brffs_smart2010_df %>%
  filter(year %in% c(2002, 2010)) %>% 
  group_by(year, locationabbr) %>%
  summarize(n_obs = n()) %>%
  filter(n_obs >= 7) %>%
  arrange(year, desc(n_obs))
```

```{r}
excellent_data =
  brfss_smart2010 %>%
  filter(Response %in% c("Excellent")) %>%
  select(Year, Locationabbr, Data_value) %>%
  group_by(Locationabbr, Year) %>%
  mutate(avg_data_value = mean(Data_value)) %>%
  select(Year, Locationabbr, avg_data_value) %>%
  distinct()
```

  * In 2002, PA, MA, NJ, CT, FL, NC (total of 6 states) were observed at 7 or more locations. In 2010, total of 14 states were observed at 7 or more locations, including FL, NJ, TX, CA, MD, NC, NE, WA, MA, NY, OH, CO, PA, SC.
  * WA were observed at 7 or more locations in 2010.
## Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state.


## Making a "spaghetti" plot
```{r}
excellent_data %>%
  ggplot(aes(x = Year, y = avg_data_value, color = Locationabbr)) +
  geom_line() +
  labs(title = "Average Value of 'Excellent' Responses Over Time", x = "Year", y = "Average Value") +
  theme_minimal()
```

  * Made a plot showing a line for each state of the United States across years.
  * Used the `geom_line` geometry and `group` aesthetic functions.
  * The spaghetti plot of `Average Value of Excellent Responses Over Time` includes colored lines of 50 states, `x-axis` of year from 2002 to 2010 and `y-axis` of average value.
  * According to this plot, the `Excellent` responses seem to be clustered the most during 2004 and 2008.
 
 
## Making a two-panel plot
```{r}
brfss_smart2010 %>%
  select(Year, Data_value, Locationabbr, Locationdesc, Response) %>%
  filter(
    Year %in% c(2006, 2010),
    Locationabbr == "NY") %>% 
  ggplot(aes(x = Data_value, fill = Response)) +
  geom_boxplot(alpha = 0.5) +
  labs(title = "Distribution of Data_value in NY State (2006 & 2010)",
       x = "Data Value", y = "Response") +
  theme_minimal() +
  facet_grid(. ~ Year)
```

  * Used `select` function to select the variables from the dataset.
  * Filtered specfic years of 2006 and 2010 by using `filter` function.
  * Specified the location to NY State.
  * Made a two-panel, box plot for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.
  * To create the two-panel, boxplot, `ggplot`, and `geom_point` were used.
  * Added the title by using `labs` function.
  * Used `theme_minimal` function for the background.
  * Finally, using `facet_grid` function, plots for 2006 and 2020 were put together side to side. 
  * Even though I filtered the responses from “Excellent” to “Poor” above using `filter` and `mutate` function, I kept getting the same boxplots that include every response. I tried several different ways to filter responses and create the accurate plot, I couldn't make the one with only 5 responses of Excellent to Poor. 



## Problem 3

```{r}
library(dplyr)
```

  * Loaded `library(dplyr)` for the further use. 

## Load, tidy, merge and organize the data sets.
```{r}
covar_df= 
  read.csv("~/Desktop/Hw3/nhanes_covar.csv", skip = 4) %>%
  janitor::clean_names() %>%
  drop_na() %>%
  filter(age >= 21) %>% 
  mutate(education = recode(education, "1" = "Less than high school", "2"= "High school equivalent", "3" = "More than high school")) %>%
  mutate(
    sex = recode(sex, "1" = "male", "2" = "female"),
    education = recode(education, "1" = "male")
  )

accel_df =
  read.csv("~/Desktop/Hw3/nhanes_accel.csv") %>%
  janitor::clean_names()

joined_data =
  accel_df %>%
  inner_join(covar_df, by = "seqn") %>%
  relocate(seqn, sex, age, bmi, education) %>%
  filter(age >= 21) %>%
  mutate(
    sex = forcats::fct_relevel(sex, "male", "female"),
    education = forcats::fct_relevel(education, "Less than high school", "High school equivalent", "More than high school")
  )
view(joined_data)
```

  * Loaded the csv file of each data set by using `read.csv`.
  * Tidyed the data set by cleaning the name (`janitor::clean_names` was used), dropped unnecessary data using `drop_na` function. 
  * Then, filtered to exclude participants with less than 21 years of age and the gender was defined, other than using the numeric values of 1 and 2. `mutate` and `recode` functions were used.
  * Used `recode` again for defining and filtering the education variable as well.
  * For accel data set, same processes were repeated for filtering age.
  * `inner_join` function and `mutate` functions were used in advance to merge two different data sets. (Organized the accel data set in order of seqn, sex, age, bmi, and education. Education level and sex categories were organized by following the `forcats::fct_relevel` functions.)
  * Merged two organzied data sets.
  
  
## Creating a table for the number of men and women in each category.
```{r}
education_table =
  joined_data %>%
  group_by(education, sex) %>%
  summarize(count = n())

knitr::kable(education_table, caption = "Number of Men and Women in each Education Category", format = "markdown")
```

  * Created a table using the merged data.
  * Grouped the data by education and sex variables.
  * Then to load the table, used `kintr::kable` function.
  * The table provides the participant distribution based on their education level and gender. It offers insights into the number of participants in each education category, segmented by their gender. There are more females (28) than males (27) in the "Less Than High School Education" category. "High School Equivalent" exhibits a higher representation of males (35) compared to females (23). In the "More Than High School" category, threr are slightly more females (59) than males (56).
  
  
## Creating a visualization of the age distributions for men and women in each education category.
```{r}
age_distribition_viz =
  ggplot(joined_data, aes(x = education, y = age, fill = sex)) +
  geom_boxplot() +
  labs(title = "Age Distribution by Gender and Education",
       x = "Education Level",
       y = "Age") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 40, hjust =1))
```

  * The box plot contains x-axis of `Education Level`: `Less than high school`, `High school equivalent`, and `More than high school` and y-axis of age from 20~ 80.
  * Male is colored purple, and female is colored yellow in the box plots.
  * For `Less than high shcool` category, IQR ranges are from early 40s to around 70s for both genders.
  * For `High school equivalent` category, the minimum IQR of male starts earlier (mid 30s) and maximum IQR ends around mid 60s. For women, the minimum IQR starts around mid 40s and the maximum IQR ends around mid 70s.Lastly, `More than high school` category, minimum IQR for both genders start around early 30s to around early 60s.
  * In the `Less Than High School` category, age distributions are wide for both genders, with a slightly older among women. In `High School Equivalent`, both genders have broad age distributions, but women's age is slightly older. In "More Than High School", women and men both start at a younger age, indicating earlier completion of higher education.

 
  
## Plotting for the total activities that compare men to women.
```{r}
activity_total =
joined_data %>%
  rowwise() %>%
  mutate(
    total_activity = sum(c_across(min1:min1440)) %>%
      as.numeric())

total_activity_plot =
  ggplot(activity_total, aes(x = age, y = total_activity, color = sex)) +
  geom_point(alpha = 0.5) +
  facet_wrap(~ education) +
  geom_smooth() +
  labs(title = "Total Activity by Age, Genderm and Education Level",
       x = "Age",
       y = "Total Activity",
       color = "Gender") +
  theme_minimal()
```

  * Based on the scatter plot depicting total activity by age, sex, and education level, several observations can be made
  * Total activity tends to decrease with age across all education categories for both genders.
  * Women exhibit higher total activity levels than men in younger age groups, particularly in the `less than high school` category.
  * Notably, the `Less than high school` category demonstrates elevated total activity for both genders, particularly in the younger age groups.
  * The steepness of the trend lines on the graph indicates that individuals with lower education levels tend to experience a more pronounced decline in total activity as they age. In particular, the `Less than high school` category shows a steeper decrease in total activity with increasing age compared to other education categories, suggesting that those with less education may be more vulnerable to reduced physical activity as they grow older.
  
  
  
## Making a three-panel plot for the 24-hour activity time courses.
```{r}
activity_data =
  joined_data %>%
  pivot_longer(cols = starts_with("min"), names_to = "minute", values_to = "activity") %>%
  mutate(
    minute = as.numeric(gsub("min", "", minute)),
    hour = (minute - 1) %/% 60) %>%
  group_by(education, sex, hour) %>%
  summarize(mean_activity = mean(activity)) %>%
  ungroup()

activity_data %>%
  ggplot(aes(x = hour, y = mean_activity, color = sex)) +
  geom_line() +
  geom_smooth(method = "lm", se = FALSE, size = 1) +
  facet_wrap(. ~ education, scales = "free") +
  labs(title = "24-Hour Activity by Education and Gender",
       x = "Hour of the Day",
       y = "Mean Value of Activity") +
  theme_minimal()
```


  * Based on the findings from the three-panel plot displaying the 24-hour activity patterns categorized by education level and gender, several observations come to light:
  * Across all education levels, males tend to exhibit higher mean activity levels compared to females at every hour throughout the day, with one notable exception: the `Less than high school` group, where the trend line for female mean activity surpasses that of males during the late afternoon and early evening hours.
  * Individuals with `More than a high school` education consistently demonstrate higher mean activity levels throughout the day when contrasted with those with lower education levels.
  * On a general note, the data suggests that mean activity levels are at their peak in the morning, particularly between 8 am and 10 am, and gradually decrease as the day progresses. The lowest levels are observed during the evening, typically between 8 pm and 10 pm.







